Bayesian methods for black hole merger parameter estimation with LISA.

### 2019-03-13 paper/paper_PE.tex
Some initial text for SOBH section

\newcommand{\SM}[1]{{\color{Blue} #1}}
\newcommand{\jgb}[1]{{\color{DarkGreen} #1}}

@@

>>>>

\newcommand{\SM}[1]{{\color{Blue} #1}}
\newcommand{\jgb}[1]{{\color{DarkGreen} #1}}
\newcommand{\tdc}[1]{{\color{red} #1}}

@@

Stellar-origin black-hole (SOBH) mergers have been recently recognized as an
important source of LISA detections \cite{Sesana16} although full parameter
estimation studies are not available yet. In this section we present
three case studies of parameter inference for such sources. The parameters of
the simulated mergers are taken from the Radler LISA data challenge and are
listed in Table \ref{table:sobh_params}.
\begin{table}
	\begin{tabular}{|c||c|c|c|}
		\hline
		Radler entry 		& 1 	& 12 	& 52 \\
		\hline
		Mass 1 ($\Msol$) 	& 21.4	& 63.5	& 57.0 \\
		\hline
		Mass 2 ($\Msol$) 	& 20.1	& 45.1	& 36.8 \\
		\hline
		Distance (Mpc) 		& 49.1	& 318	& 308 \\
		\hline
		LISA SNR 			& 26.6	& 11.9	& 16.0 \\
		\hline
	\end{tabular}
	\caption{Parameters of the simulated SOBH mergers. \tdc{Maybe we can show
			 these on the upper right of each corner plot instead.}}
	\label{table:sobh_params}
\end{table}

The waveform model used to work with such systems is the phenomenological
approximant IMRPhenomD, extended at low frequency with a post-Newtonian
frequency-domain waveform. In practice, because LISA only observes the early
inspiral of SOBH systems, only the post-Newtonian waveform plays a role here.
We do not expect radiation modes beyond the quadrupole to give a significant
contribution to the parameter estimates, so we only include the $(2,2)$ mode
for simplicity. We do not include any effect arising from black-hole spins.

As done in the previous section, we apply our parameter inference procedure to
these systems using both BAMBI and parallel-tempered MCMC sampling, and we
numerically evaluate the Fisher matrix at the parameters of the injection.

Due to the duration of an SOBH inspiral in the LISA band, and the complete lack
of merger and ringdown, the chirp mass $\Mchirp$ is measured much better than
other mass parameters and the component masses have a much stronger correlation
than usually found with ground-based gravitational-wave detectors. In order for
the sampling to converge robustly in a reasonable time, we find it is necessary
to express the mass parameters in terms of $\Mchirp$ and symmetric mass ratio
$\eta$, and a relatively tight prior around the injected parameters is required
as well.

In order to generate these results, our PTMCMC sampler requires between
$5\times 10^5$ and $3\times 10^6$ steps with 100 temperatures, taking on the
order of days on a single CPU core without careful optimization. The BAMBI
sampling, on the other hand, completes in tens of minutes only.

The resulting posterior distributions are shown in Fig.~\ref{fig:sobh_corner_1}
and \ref{fig:sobh_corner_12}. We find close agreement between the BAMBI and
PTMCMC samplers. The full joint distributions are fairly complex, but the
complexity is mostly restricted to multimodalities and degeneracies in the
angular parameters only. The typical degeneracy between inclination and distance
is evident. Masses, merger time, distance, inclination and sky location, which
are arguably the most astrophysically useful parameters, have simple marginal
distributions with a well-defined mode and close to Gaussian, even at these
moderate SNR values. The Fisher matrix estimates represent the marginal
distributions of these parameters at least to the order of magnitude. In some
cases the Fisher estimates approximate very precisely even the joint
distributions of pairs of these parameters.

\begin{figure*}
	\includegraphics[width=\textwidth]{../plots/corner_sobh_tdc1_ptmcmc_bambi}
	\caption{Inferred parameter posterior distribution for SOBH system 1.}
	\label{fig:sobh_corner_1}
\end{figure*}

\begin{figure*}
	\includegraphics[width=\textwidth]{../plots/corner_sobh_tdc2_ptmcmc_bambi}
	\caption{Inferred parameter posterior distribution for SOBH system 12.}
	\label{fig:sobh_corner_12}
\end{figure*}
